---
title: Data Vis Chapter 6
author: wyih
date: '2019-09-26'
slug: data-vis-chapter-6
categories:
  - R
tags:
  - socviz
subtitle: ''
summary: ''
authors: []
lastmod: '2019-09-26T16:21:48+08:00'
featured: no
output:
  blogdown::html_page:
    toc: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---
```{r setup, include=FALSE}
library(tidyverse)
library(socviz)
library(gapminder)
```

```{r}
p <-  ggplot(data = gapminder,
             mapping = aes(x = log(gdpPercap), y = lifeExp))

p + geom_point(alpha = 0.1) +
  geom_smooth(color = "tomato",
              fill = "tomato",
              method = MASS::rlm) + #robust regression line
  geom_smooth(color = "steelblue",
              fill = "steelblue",
              method = "lm")

p + geom_point(alpha = 0.1) +
  geom_smooth(
    color = "tomato",
    method = "lm",
    size = 1.2,
    formula = y ~ splines::bs(x, 3),
    se = FALSE
  )

p + geom_point(alpha = 0.1) +
  geom_quantile( # specialized version of geom)smooth that can fit quantile regression
    color = "tomato",
    size = 1.2,
    method = "rqss",
    lambda = 1,
    quantiles = c(0.20, 0.5, 0.85)
  )
```

## Show Several Fits at Once, with a Legend
```{r}
model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors

p0 <- ggplot(data = gapminder,
             mapping = aes(x = log(gdpPercap), y = lifeExp))

p1 <- p0 + geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) +
  geom_smooth(
    method = "lm",
    formula = y ~ splines::bs(x, df = 3),
    aes(color = "Cubic Spline", fill = "Cubic Spline")
  ) +
  geom_smooth(method = "loess",
              aes(color = "LOESS", fill = "LOESS"))

p1 + scale_color_manual(name = "Models", values = model_colors) +
  scale_fill_manual(name = "Models", values = model_colors) +
  theme(legend.position = "top")
```

## Model-based Graphics
```{r}
min_gdp <- min(gapminder$gdpPercap)
max_gdp <- max(gapminder$gdpPercap)
med_pop <- median(gapminder$pop)

pred_df <- expand.grid(gdpPercap = (seq(from = min_gdp, to = max_gdp,
length.out = 100)), pop = med_pop, continent = c("Africa",
"Americas", "Asia", "Europe", "Oceania"))

dim(pred_df)

out <- lm(formula = lifeExp ~ gdpPercap + pop + continent, data = gapminder)

pred_out <- predict(object = out, newdata = pred_df, interval = "predict")
pred_df <- cbind(pred_df, pred_out)

```
```{r}
p <-
  ggplot(
    data = subset(pred_df, continent %in% c("Europe", "Africa")),
    aes(
      x = gdpPercap,
      y = fit,
      ymin = lwr,
      ymax = upr,
      color = continent,
      fill = continent,
      group = continent
    )
  )

p + geom_point(
  data = subset(gapminder,
                continent %in% c("Europe", "Africa")),
  aes(x = gdpPercap, y = lifeExp,
      color = continent),
  alpha = 0.5,
  inherit.aes = FALSE
) +
  geom_line() +
  geom_ribbon(alpha = 0.2, color = FALSE) +
  scale_x_log10(labels = scales::dollar)
```

## Tidy Model Objects with Broom

### get component-level statistics with tidy()
```{r}
library(broom)
out_comp <- tidy(out)
out_comp %>% round_df()
```
"not in" `%nin%` is availabe via `socviz`.
`prefix_strip` from `socviz` drops prefixes
```{r}
#confidence interval
out_conf <- tidy(out, conf.int = TRUE)
out_conf <- subset(out_conf, term %nin% "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")
```
```{r}
p <- ggplot(out_conf,
            mapping = aes(
              x = reorder(nicelabs, estimate),
              y = estimate,
              ymin = conf.low,
              ymax = conf.high
            ))
p + geom_pointrange() + coord_flip() + labs(x = "", y = "OLS Estimate")
```

### Get observation-level statistics with augment()

```{r}
out_aug <- augment(out)
p <- ggplot(data = out_aug, mapping = aes(x = .fitted, y = .resid))
p + geom_point()
```
### Get model-level statistics with glance()
Broom is able to `tidy` (and `augment`, and `glance` at) a wide range of model types. 
```{r}
library(survival)

out_cph <- coxph(Surv(time, status) ~ age + sex, data = lung)
out_surv <- survfit(out_cph)
out_tidy <- tidy(out_surv)
p <- ggplot(data = out_tidy, mapping = aes(time, estimate))
p + geom_line() + geom_ribbon(mapping = aes(ymin = conf.low,
                                            ymax = conf.high),
                              alpha = 0.2)
```

## Grouped Analysis
`nest` and `unnest`
```{r}
out_le <- gapminder %>%
  group_by(continent, year) %>%
  nest()
```
```{r}
fit_ols <- function(df) {
  lm(lifeExp ~ log(gdpPercap), data = df)
}

out_le <- gapminder %>%
  group_by(continent, year) %>%
  nest() %>%
  mutate(model = map(data, fit_ols))



out_tidy <- gapminder %>%
  group_by(continent, year) %>%
  nest() %>%
  mutate(model = map(data, fit_ols),
         tidied = map(model, tidy)) %>%
  unnest(tidied, .drop = TRUE) %>%
  filter(term %nin% "(Intercept)" &
           continent %nin% "Oceania")

p <- ggplot(
  data = out_tidy,
  mapping = aes(
    x = year,
    y = estimate,
    ymin = estimate - 2 * std.error,
    ymax = estimate + 2 * std.error,
    group = continent,
    color = continent
  )
)

p + geom_pointrange(position = position_dodge(width = 1)) +
  scale_x_continuous(breaks = unique(gapminder$year)) +
  theme(legend.position = "top") +
  labs(x = "Year", y = "Estimate", color = "Continent")
```
## Plot Marginal Effects
```{r}
library(margins)
gss_sm$polviews_m <- relevel(gss_sm$polviews, ref = "Moderate")
out_bo <- glm(obama ~ polviews_m + sex * race,
              family = "binomial",
              data = gss_sm)
summary(out_bo)
bo_m <- margins(out_bo)
summary(bo_m)
```
The margins library comes with several plot methods of its own. If you wish, at this point you can just try `plot(bo_m)` to see a plot of the average marginal effects, produced with the general look of a Stata graphic. Other plot methods in the margins
library include `cplot()`, which visualizes marginal effects conditional on a second variable, and `image()`, which shows predictions or marginal effects as a filled heatmap or contour plot.

```{r}
bo_gg <- as_tibble(summary(bo_m))
prefixes <- c("polviews_m", "sex")
bo_gg$factor <- prefix_strip(bo_gg$factor, prefixes)
bo_gg$factor <- prefix_replace(bo_gg$factor, "race", "Race: ")

bo_gg %>% select(factor, AME, lower, upper)

p <- ggplot(data = bo_gg, aes(
  x = reorder(factor, AME),
  y = AME,
  ymin = lower,
  ymax = upper
))

p + geom_hline(yintercept = 0, color = "gray80") +
  geom_pointrange() + coord_flip() +
  labs(x = NULL, y = "Average Marginal Effect")
```
```{r}
pv_cp <- cplot(out_bo, x = "sex", draw = FALSE)

p <- ggplot(data = pv_cp, aes(
  x = reorder(xvals, yvals),
  y = yvals,
  ymin = lower,
  ymax = upper
))

p + geom_hline(yintercept = 0, color = "gray80") +
  geom_pointrange() + coord_flip() +
  labs(x = NULL, y = "Conditional Effect")
```

## Plots for Surveys
```{r}
library(survey)
library(srvyr)

options(survey.lonely.psu = "adjust")
options(na.action = "na.pass")

gss_wt <- subset(gss_lon, year > 1974) %>%
  mutate(stratvar = interaction(year, vstrat)) %>%
  as_survey_design(
    ids = vpsu,
    strata = stratvar,
    weights = wtssall,
    nest = TRUE
  )

out_grp <- gss_wt %>%
  filter(year %in% seq(1976, 2016, by = 4)) %>%
  group_by(year, race, degree) %>%
  summarize(prop = survey_mean(na.rm = TRUE)) # calculate  properly weighted survey means

out_mrg <- gss_wt %>%
  filter(year %in% seq(1976, 2016, by = 4)) %>%
  mutate(racedeg = interaction(race, degree)) %>%
  group_by(year, racedeg) %>%
  summarize(prop = survey_mean(na.rm = TRUE))


out_mrg <- gss_wt %>%  filter(year %in% seq(1976, 2016, by = 4)) %>%
  mutate(racedeg = interaction(race, degree)) %>% group_by(year,
                                                           racedeg) %>% 
  summarize(prop = survey_mean(na.rm = TRUE)) %>%
  separate(racedeg, sep = "\\.", into = c("race", "degree")) 

p <- ggplot(
  data = subset(out_grp, race %nin% "Other"),
  mapping = aes(
    x = degree,
    y = prop,
    ymin = prop - 2 * prop_se,
    ymax = prop + 2 * prop_se,
    fill = race,
    color = race,
    group = race
  )
)

dodge <- position_dodge(width = 0.9)

p + geom_col(position = dodge, alpha = 0.2) +
  geom_errorbar(position = dodge, width = 0.2) +
  scale_x_discrete(labels = scales::wrap_format(10)) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  labs(
    title = "Educational Attainment by Race",
    subtitle = "GSS 1976-2016",
    fill = "Race",
    color = "Race",
    x = NULL,
    y = "Percent"
  ) +
  facet_wrap( ~ year, ncol = 2) +
  theme(legend.position = "top")

```
```{r}
p <- ggplot(
  data = subset(out_grp, race %nin% "Other"),
  mapping = aes(
    x = year,
    y = prop,
    ymin = prop - 2 * prop_se,
    ymax = prop + 2 * prop_se,
    fill = race,
    color = race,
    group = race
  )
)

p + geom_ribbon(alpha = 0.3, aes(color = NULL)) + #Use ribbon to show the error range
  geom_line() + #Use line to show a time trend
  facet_wrap( ~ degree, ncol = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  labs(
    title = "Educational Attainment by Race",
    subtitle = "GSS 1976-2016",
    fill = "Race",
    color = "Race",
    x = NULL,
    y = "Percent"
  ) +
  theme(legend.position = "top")
```

Other useful packages: `infer`, `ggally`