---
title: Data Visualization Chapter 2-4
author: wyih
date: '2019-09-26'
slug: test
categories:
  - R
  - Books
tags:
  - socviz
subtitle: ''
summary: ''
authors: []
lastmod: '2019-09-26T00:02:42+08:00'
output:
  blogdown::html_page:
    toc: true
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
library(tidyverse)
library(socviz)
library(gapminder)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)
```

## Chapter 2
`geom_point`
```{r cars}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point()
```

## Chapter 3
`geom_smooth`

```{r pressure, echo=FALSE}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth()
```

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth()
```

`scale_x_log10`

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth(method = "gam") + scale_x_log10()
```

`scales::dollar`
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() +
geom_smooth(method = "gam") +
scale_x_log10(labels = scales::dollar)
```

### Wrong way to set color
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp,
color = "purple"))
p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()
```

The `aes()` function is for mappings only. Do not use it to change properties to a particular value. If we want to set a property, we do it in the geom_ we are using, and outside the `mapping =aes(...) `step. 
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(color = "purple") + geom_smooth(method = "loess") + scale_x_log10()
```
The various `geom_` functions can take many other arguments that will affect how the plot looks but do not involve mapping variables to aesthetic elements.
 “alpha” is an aesthetic property that points (and some other plot elements) have, and to which variables can be mapped. It controls how transparent the object will appear when drawn. It’s measured on a scale of zero to one. 

```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(alpha = 0.3) + geom_smooth(color = "orange", se = FALSE,
                                          size = 8, method = "lm") + scale_x_log10()
```
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y=lifeExp))
p + geom_point(alpha = 0.3) +
  geom_smooth(method = "gam") +
  scale_x_log10(labels = scales::dollar) +
  labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
       title = "Economic Growth and Life Expectancy",
       subtitle = "Data points are country-years",
       caption = "Source: Gapminder.")
```
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp,
                                            color = continent))
p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()
```
The color of the standard error ribbon is controlled by the fill aesthetic. 
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp,
                                            color = continent, fill = continent))
p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()
```

### Aesthetics Can Be Mapped per Geom
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(mapping = aes(color = factor(year))) + 
  geom_smooth(method = "loess") +
  scale_x_log10()
```

Order doesn't matter!!!
Besides `scale_x_log10()`, you can try `scale_x_sqrt()` and `scale_x_reverse()`
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = pop, y = lifeExp))
p + geom_smooth(method = "loess") + 
  geom_point(mapping = aes(color = continent)) + 
  scale_x_reverse(labels = scales::number)
```
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(mapping = aes(color = log(pop))) + scale_x_log10()
```

### Save plots
```{r ,eval=FALSE}
p_out <-  p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()
ggsave("my_figure.pdf", plot = p_out)
```

## Chapter 4
### Group data and the "Group" Aesthetic
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line()
```
use the `group` aesthetic to tell ggplot explicitly about this country-level structure
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(aes(group = country))
```

### Facet to make small multiples
 use `facet_wrap()` to split our plot by `continent`.
 
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(aes(group = country)) + facet_wrap(~continent)
```
Add another enhancements
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(color="gray70", aes(group = country)) + 
  geom_smooth(size= 1.1, method = "loess", se = FALSE) +
  scale_y_log10(labels=scales::dollar) +
  facet_wrap(~continent , ncol = 5) +
  labs(x = "Year",
       y = "GDP per capita on Five Continents")
```

Use `facet_grid`
```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) +
  geom_smooth() + 
  facet_grid(sex ~ race)
```
```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) +
  geom_smooth() + 
  facet_grid(sex ~ race + degree)
```

### Geoms can transform data

```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar()
```

`geom_bar` called the default `stat_` function associated with it,`stat_count()`.

```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop..))
```

```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop.., group = 1))
```

```{r}
table(gss_sm$religion)
```

```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = religion, color = religion))
p + geom_bar()

p <- ggplot(data = gss_sm, mapping = aes(x = religion, fill = religion))
p + geom_bar() + guides(fill = FALSE)
p + geom_bar()
```

```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar()

```
```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "fill")

```

if you want separate bars
```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "dodge", mapping = aes(y = ..prop..,
                                               group = religion))
```
However, they don't sum to one within each region. They sum to one across regions.
```{r}
p <- ggplot(data = gss_sm, mapping = aes(x = religion))
p + geom_bar(position = "dodge", mapping = aes(y = ..prop..,
                                               group = bigregion)) +
  facet_wrap(~bigregion, ncol=1)
```

### Histgrams and Density Plots
```{r}
p <- ggplot(data = midwest, mapping = aes( x = area))
p + geom_histogram()
```

```{r}
p <- ggplot(data = midwest, mapping = aes( x = area))
p + geom_histogram(bins = 10)
```

```{r}
oh_wi <- c("OH", "WI")
p <- ggplot(data = subset(midwest, subset = state %in% oh_wi),
            mapping = aes(x = percollege, fill = state))
p + geom_histogram(alpha = 0.4, bins = 20)
```

```{r}
p <- ggplot(data = midwest, mapping = aes( x = area))
p + geom_density()
```
```{r}
p <- ggplot(data = midwest, mapping = aes( x = area, fill = state,
                                           color = state))
p + geom_density(alpha = 0.3)
```

### Avoid Transformations When Necessary

```{r}
p <- ggplot(data = titanic, mapping = aes(x = fate, y = percent,
                                          fill = sex))
p + geom_bar(position = "dodge", stat = "identity") + theme(legend.position = "top")
```
```{r}
p <- ggplot(data = oecd_sum,
            mapping = aes(x = year, y = diff, fill = hi_lo))
p + geom_col() + guides(fill = FALSE) + 
  labs(x = NULL, y = "Difference in Years",
       title = "The US Life Expectancy Gap",
       subtitle = "Difference between US and OECD 
       average life expectancies, 1960-2015",
       caption = "Data: OECD. After a chart by Christopher Ingraham,
       Washington Post, December 27th 2017.")
```

